Fast solvers for the high-order FEM simplicial de Rham complex \thanks{Submitted to the editors July XX, 2024. \funding{ PDB and PEF were supported by EPSRC grant EP/W026260/1. } }

### 2024-06-27 paper/paper.tex
Writing with Pablo

\newcommand{\pef}[1]{\todo[inline,color=blue!40]{PEF: #1}}

@@

We discretize the $\Ltwo$ de Rham complex on triangular and tetrahedral meshes with finite element subcomplexes of the first and second kinds.

\begin{figure}[htbp] 
\centering
\begin{tikzcd}
  \CG_p \arrow[r, "\grad"] & \Ned_p \arrow[r, "\curl"] & 
   \RT_p \arrow[r, "\div"] & \DG_{p-1} \\
  \Hgrad \arrow[r, "\grad"] \arrow[u] \arrow[d]  & \Hcurl \arrow[r, "\curl"]
  \arrow[d] \arrow[u] & \Hdiv \arrow[r, "\div"] \arrow[u] \arrow[d] & \Ltwo \arrow[u] \arrow[d] \\
   \CG_p \arrow[r, "\grad"] & \NedTwo_{p-1} \arrow[r, "\curl"] & 
   \BDM_{p-2} \arrow[r, "\div"] & \DG_{p-3}
\end{tikzcd}
\caption{The $\Ltwo$ de Rham complex (middle), and the finite element subcomplexes of the first (above) and second kinds (below).}
\end{figure}
\pef{Fix vertical alignment}

For each finite element subcomplex built on a mesh $\mesh$ of a domain $\Omega \subset \mathbb{R}^d$, we denote the spaces discretizing $k$-forms $H\Lambda^k(\Omega)$ with polynomial degree $p$ as $X^k_p(\mesh)$.
These are built from cellwise spaces via the standard pullback construction from a reference cell $\Khat$:
\begin{equation}
X^k_{h,p} = X^k_p(\mesh) = \{ v \in H\Lambda^k(\Omega): \Forall K \in \mesh \Exists \hat{v} \in X^k_p(\Khat) \text{ s.t.} \left.v\right|_K = \pullback^k_K(\hat{v}) \},
\end{equation}
where $\pullback^k_K$ denotes the appropriate pullback that preserves continuity of the traces of a $k$-form across cell facets.

We follow the definition of interpolation degrees of freedom from
Demkowicz et al.~\cite{demkowicz00} to construct dual bases that promote orthogonality in the
$\Ltwo(\Khat)$ and $H(\mathrm{d}, \Khat)$ inner products.
In the construction of the degrees of freedom we will employ bubble subspaces on a given mesh entity $S \subset \Khat$ with zero trace, defined as
\begin{equation}
B^k_p(S) \coloneqq \{v \in X^k_p(S): \tr v = 0 \text{ on } \partial S\}.
\end{equation}

Throughout we choose $\Khat$ to be an equilateral simplex, as this ...

@@

On $\Khat$ we discretize $\Hgrad$ with the space
\begin{equation}
X^0_{p}(\Khat) = \P_p(\Khat),
\end{equation}
where $\P_p(\Khat)$ is the space of polynomials of total degree less than or equal to $p$.
We define a basis for the dual of $X^0_p(\Khat)$ as
point evaluations at $V \in \text{vertices}(\Khat)$,
\begin{equation}
   \ell^V(v) = v(V),
\end{equation}
and for each sub-entity $S \in \text{edges}(\Khat) \union \text{faces}(\Khat) \union \text{interior}(\Khat)$
we take integral moments of surface gradients, 
\begin{equation}
   \ell^S_j(v) = (\grad_S\phi^S_j, \grad_S v)_{S},
\end{equation}
where $\grad_S$ is the tangential gradient on $S$,
and $\{\phi^S_j\}$ is a basis for $B^0_p(S)$
such that
\begin{equation} \label{eq:hgrad-fdm}
   (\grad_S\phi^S_j, \grad_S\phi^S_i)_{S} = \delta_{ij}, \quad
   (\phi^S_j, \phi^S_i)_{S} = \lambda_j\delta_{ij}.
\end{equation}
The eigenbases $\{\phi^S_j\}$ are numerically computed offline and only once 
on the reference interval, triangle, and tetrahedron.

@@

Recall the definition of the $\Hcurl$-conforming finite element spaces
\begin{equation}
   X^1_{p, p_E}(K) = \{v\in [\P_p(K)]^d : \tr v \in \P_{p_E}(E) \Forall E \in \text{edges}(K)  \}
\end{equation}
for $p_E=p-1$ we obtain the \Nedelec space of the first kind $\Ned_p(K)$, while
$p_E=p$ gives the \Nedelec space of the second kind $\NedTwo_p(K) = [\P_p(K)]^d$. 
These spaces enforce continuity of the tangential trace across mesh interfaces.

We define a basis for the dual of $X^1_{p, p_E}(K)$ as 
tangential moments along $E\in \text{edges}(\Khat)$,
\begin{equation}
   \ell^E_j(v) = (q_j, v\cdot \bt)_E, \quad q_j \in \P_{p_E}(E), 
\end{equation}
and for each sub-entity $S \in \text{faces}(\Khat) \cup \text{interior}(\Khat)$,
\begin{align}
   \ell^{S,0}_j(v) &= (\grad_S\phi^S_j, v)_{S}, \\
   \ell^{S,1}_j(v) &= (\curl_S\Phi^S_j, \curl_S v)_{S}, \label{eq:hcurl-interior-dofs}
\end{align}
where $\curl_S$ is the tangential curl on $S$,
$\{\phi^S_j\}$ are the $B^0_{p_S}(S)$ bases constructed in \eqref{eq:hgrad-fdm},
and $\{\curl_S\Phi^S_j\}$ is a basis for $\curl_S B^1_{p_S}(S)$
such that
\begin{equation} \label{eq:hcurl-fdm}
   (\curl_S\Phi^S_j, \curl_S\Phi^S_i)_{S} = \delta_{ij}, \quad
   (\Phi^S_j, \Phi^S_i)_{S} = \lambda_j\delta_{ij}.
\end{equation}

>>>>

\newcommand{\pef}[1]{\todo[inline,color=blue!40]{PEF: #1}}
\newcommand{\pbm}[1]{\todo[inline,color=red!40]{PEF: #1}}

@@

We discretize the $\Ltwo$ de Rham complex on triangular and tetrahedral meshes with finite element subcomplexes of the first and second kinds (\Cref{fig:3DL2deRham}).

\begin{figure}[htbp] 
\centering
\begin{tikzcd}
  \CG_p \arrow[r, "\grad"] & \Ned_p \arrow[r, "\curl"] & 
   \RT_p \arrow[r, "\div"] & \DG_{p-1} \\
  \Hgrad \arrow[r, "\grad"] \arrow[u] \arrow[d]  & \Hcurl \arrow[r, "\curl"]
  \arrow[d] \arrow[u] & \Hdiv \arrow[r, "\div"] \arrow[u] \arrow[d] & \Ltwo \arrow[u] \arrow[d] \\
   \CG_p \arrow[r, "\grad"] & \NedTwo_{p-1} \arrow[r, "\curl"] & 
   \BDM_{p-2} \arrow[r, "\div"] & \DG_{p-3}
\end{tikzcd}
\caption{The $\Ltwo$ de Rham complex in three dimensions (middle), and the finite element subcomplexes of the first (above) and second kinds (below).}
\label{fig:3DL2deRham}
\end{figure}
\pef{Fix vertical alignment}

For each finite element subcomplex built on a mesh $\mesh$ of a bounded Lipschitz domain $\Omega \subset \mathbb{R}^d$, we denote the spaces discretizing $k$-forms $H\Lambda^k(\Omega)$ with polynomial degree $p$ as $X^k_p(\mesh)$. We employ the same notation for spaces of the first or second kind where the distinction does not matter.
These spaces are built from cellwise spaces $X^k_p(\Khat)$ (constructed below) via the standard pullback construction from a reference cell $\Khat$:
\begin{equation}
X^k_{h,p} = X^k_p(\mesh) = \{ v \in H\Lambda^k(\Omega): \Forall K \in \mesh \Exists \hat{v} \in X^k_p(\Khat) \text{ s.t.} \left.v\right|_K = \pullback^k_K(\hat{v}) \},
\end{equation}
where $\pullback^k_K$ denotes the appropriate pullback that preserves continuity of the traces of a $k$-form across cell facets.

We follow the definition of interpolation degrees of freedom from
Demkowicz et al.~\cite{demkowicz00} to construct dual bases that promote orthogonality in the
$\Ltwo(\Khat)$ and $H(\mathrm{d}, \Khat)$ inner products.
In the construction of the degrees of freedom we will employ bubble subspaces on a given mesh entity $S \subset \Khat$ with zero trace, defined as
\begin{equation}
B^k_p(S) \coloneqq \{\left.v\right|_S \in X^k_p(\Khat): \tr \left(\left.v\right|_S\right) = 0 \text{ on } \partial S\}.
\end{equation}

Throughout we choose $\Khat$ to be an equilateral simplex, as this simplifies the construction of degrees of freedom on facet entities.
We denote the entities in $\Khat$ of dimension $k$ by $\Delta^k(\Khat)$.

@@

On $\Khat$ we discretize $\Hgrad$ with the space
\begin{equation}
X^0_{p}(\Khat) = \P_p(\Khat),
\end{equation}
where $\P_p(\Khat)$ is the space of polynomials of total degree less than or equal to $p$.
We define a basis for the dual of $X^0_p(\Khat)$ as
point evaluations at vertices $V \in \Delta^0(\Khat)$,
\begin{equation}
   \ell^V(v) = v(V),
\end{equation}
and for each higher-dimensional sub-entity $S \in \cup_{k=1}^d \Delta^k(\Khat)$
we take integral moments of surface gradients, 
\begin{equation}
   \ell^S_j(v) = (\grad_S\phi^S_j, \grad_S v)_{S},
\end{equation}
where $\grad_S$ is the tangential gradient on $S$,
and $\{\phi^S_j\}$ is a basis for $B^0_p(S)$
such that
\begin{equation} \label{eq:hgrad-fdm}
   (\grad_S\phi^S_j, \grad_S\phi^S_i)_{S} = \delta_{ij}, \quad
   (\phi^S_j, \phi^S_i)_{S} = \lambda_j\delta_{ij}.
\end{equation}
The eigenbasis $\{\phi^S_j\}$ is numerically computed offline and only once
for each kind of reference subentity of dimension greater than or equal to one. Since $\Khat$ is symmetric, only one facet, edge, etc.~needs to be computed, and the rest are constructed by congruence.

@@

Recall the definition of the $\Hcurl$-conforming finite element spaces
\pef{Find this in a book somewhere and cite it}
\begin{equation}
   X^1_{p}(\Khat) = \{v\in [\P_p(K)]^d : \tr v \in \P_{p_E}(E) \Forall E \in \Delta^1(\Khat)  \}.
\end{equation}
Choosing $p_E=p-1$ we obtain the \Nedelec space of the first kind $\Ned_p(\Khat)$, while
choosing $p_E=p$ gives the \Nedelec space of the second kind $\NedTwo_p(K) = [\P_p(\Khat)]^d$.
\pef{Cite \Nedelec}
These spaces enforce continuity of the tangential trace across mesh interfaces.

We define a basis for the dual of $X^1_p(K)$ as 
tangential moments along edges $E\in \Delta^1(\Khat)$,
\begin{equation}
   \ell^E_j(v) = (q_j, v\cdot \bt)_E, \quad q_j \in \P_{p_E}(E),
\end{equation}
and for each higher-dimensional sub-entity $S \in \cup_{k=2}^d \Delta^k(\Khat)$,
\begin{align}
   \ell^{S,0}_j(v) &= (\grad_S\phi^S_j, v)_{S}, \\
   \ell^{S,1}_j(v) &= (\curl_S\Phi^S_j, \curl_S v)_{S}, \label{eq:hcurl-interior-dofs}
\end{align}
where $\curl_S$ is the tangential curl on $S$,
$\{\phi^S_j\}$ is the basis for $B^0_{p_S}(S)$ constructed in \eqref{eq:hgrad-fdm},
and $\{\curl_S\Phi^S_j\}$ is a basis for $\curl_S B^1_{p_S}(S)$
constructed by solving the following problem:
find $\{\Phi^S_j\} \subset B^1_{p_S}(S)$ such that
\begin{equation} \label{eq:hcurl-fdm}
   (\curl_S\Phi^S_j, \curl_S\Phi^S_i)_{S} = \delta_{ij}, \quad
   (\Phi^S_j, \Phi^S_i)_{S} = \lambda_j\delta_{ij}, \quad \lambda_j > 0.
\end{equation}
\pef{FIXME: what is $p_S$?}

The eigenbasis $\{\Phi^S_j\}$ is numerically computed offline and only once for each kind of reference subentity of dimension greater than or equal to two.
